---
title: "Accessing an API"
author: "Danny Brand"
date: "14-5-2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 5.3 Accessing an API

A REST API is simply presented as a URL, to which a specific service can be accessed, and even parameters passed (when necessary). A URL may simply look like:
`https://someserver.com/getdata`
or may be more complicated, with queries, authentication, and so on. Let’s look a simple example, using the `httr` package:

```{r api1}
library(httr)

# The first step is to reach out to the server, and get a response.
# Here, g is a complicated object that we do not inspect directly.
g <- GET("https://opendata.cbs.nl/ODataApi/OData/37789ksz/TypedDataSet")

# Important: if the server could be reached, you have permission, 
# and the service is available, you should see 200 as the response.
# Other HTML codes include 404 (service not found).
status_code(g)
```

```{r api2}
# We can read the actual data with the `content` function:
dat <- content(g)

# In this case, dat has two components - the interesting one is 'value'
uitkering_data_list <- dat$value
```

Unfortunately the result is a nested list, with NULL elements in some. To make this into a dataframe, we want to flatten each list, each of these will become a row in our dataframe, and bind all the rows together. This is such a common problem that I will show a robust solution:

```{r api3, warning=FALSE, message=FALSE, error=FALSE}
library(dplyr) # bind_rows
library(purrr) # flatten_dfr

uitkering_data <- lapply(uitkering_data_list, flatten_dfr) %>%
  bind_rows
```

We now have data from the Dutch government on social benefits, a dataset that gets updated frequently. The use of the API directly rather than downloading the data manually is clearly a much better option.

As we saw in this example, the challenge is often not so much getting something from the web service, but often we have substantial work to shape it into a useful object in R (usually: a dataframe).

## 5.3.1 Oefening
*Dutch demography data since 1898*

Haal data op van deze api: `https://opendata.cbs.nl/ODataApi/odata/37556/TypedDataSet`

```{r oefening1}
# The first step is to reach out to the server, and get a response.
getoef <- GET("https://opendata.cbs.nl/ODataApi/odata/37556/TypedDataSet")

# We can read the actual data with the `content` function:
datoef <- content(getoef)

# In this case, dat has two components - the interesting one is 'value'
demography_list <- datoef$value
```

Unfortunately the result is a nested list, with NULL elements in some. To make this into a dataframe, we want to flatten each list, each of these will become a row in our dataframe, and bind all the rows together. This is such a common problem that I will show a robust solution:

```{r oefening2, warning=FALSE, message=FALSE, error=FALSE}
library(dplyr) # bind_rows
library(purrr) # flatten_dfr

demography_data <- lapply(demography_list, flatten_dfr) %>%
  bind_rows
```

```{r oefening3}

knitr::kable(demography_data[1:5,1:8], caption = 'Dutch demography data since 1898')

```
